---
title: "case_study_gapminder_forestLand"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Forest Land Gapminder Study Case

I decided to analyze the Forest Area Dataset, found it on gapminder Geography Dataset: https://docs.google.com/spreadsheets/d/1X4_PAIzsvaiJVSIdkcO-cH2jFdzZC-XOSYIS5gnbN64/pub?gid=0

First Lets Load our Libs.
```{r Loading Library}
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
library(ggrepel) # Avoid Label Overlap

```

I hate working with .xl** files so i converted the Excel File to CSV using Apple Numbers, they use ";" as a separator, it confuses me when i imported first time without the sep param.
```{r Beginning to work on Dataset}
df <- read.csv("forest_land.csv",header = TRUE, sep = ";",stringsAsFactors = F)

summary(df)

View(df)


```

So much Na :/ , Lets just drop and perform some wrangling
```{r Working on Data}
#Found so much NA... Dropping Columns where all Values area NA
df <- df[,colSums(is.na(df))<nrow(df)]

#Just Find that this columns are chars.... fixing characters values
df$X2000 <- as.integer(df$X1990)
df$X2005 <- as.integer(df$X2005)

```

Done... Now we're safe to do some plots :slight_smile: 
```{r Plot Time}
#Gathering Columns to better plot
gathered_df <- gather(df,"Year","value",2:5)
     
#Values are large here, so i choose a log scale to minimize the problem
ggplot(gathered_df, 
       aes(x=Forest.area..sq..km., y=log10(value+1),
           colour = Year, size = value)) +
      geom_point(shape=1) +
      xlab("Country") +
      ylab("Forest Area(sq km)")
```

So... this plot is kinda... confusing... Lets try another Graphic with Facet Wrap by Year

```{r We Can do it better}
#Need some variable to Color, lets try continents
#luckly there's another dataset with continents of each country
country_continents <- read.csv("Countries-Continents.csv")

#excluding unecessary columns
country_continents <- country_continents[,1:2]

#merging
gathered_continent_df <- merge(gathered_df,country_continents, by.x = c("Forest.area..sq..km."), by.y = c("Country"))

#Better ScatterPlot
ggplot(gathered_continent_df, 
       aes(x=Forest.area..sq..km., y=log10(value+1), 
           colour = Continent, size = value, 
           label= Forest.area..sq..km.)) +
      facet_wrap(~Year) +
      geom_text_repel(size=2) +
      geom_point(shape=1) +
      xlab("Country") +
      ylab("Forest Area(sq km)")
```

Its clearer but  we Still Dont Know any difference of forest area between the Years, Trying by difference now

```{r Forest Area Difference}
#Difference between the oldest and the newest area
df$difference <- df$X2010 - df$X1990

#merging, just discovered another way
gathered_continent_df$difference <- df$difference[match(gathered_continent_df$Forest.area..sq..km.,df$Forest.area..sq..km.)]

#plotting the difference
qplot(difference, data = gathered_continent_df, 
      fill = Continent, geom = "bar", binwidth = 145000) +
      scale_x_continuous(breaks = seq(-500000,500000,145000)) +
      scale_y_continuous(limits = c(0,650), breaks = seq(0,650,50))
```

Seems like there no difference in most cases(**mode = 0**, we will see it later), seems like an **South American** country lost so many forest area and an **Asian** country gets a large area lets find it.

```{r Some Console operations to feed our curiosity}
#R dont have built-in mode, so lets implement
getmode <- function(v) {
    uniqv <- unique(v)
    uniqv[which.max(tabulate(match(v, uniqv)))]
}

summary(gathered_continent_df$difference)

getmode(gathered_continent_df$difference)

#Max Difference For Each Continent, Thanks Stack Overflow
gathered_continent_df %>% 
      group_by(Continent) %>%
      slice(which.max(difference)) 

#Now the minimum
gathered_continent_df %>% 
      group_by(Continent) %>%
      slice(which.min(difference)) 
```

Intesting, Brazil has the largest forest area but have several loss over the years
Europe is doing a good job, Armenia is not even close the second minimum.

## Additional Info

```{r Everyone Loves Bars}
 #Bar, just because everyone loves Bars, log scale can be decieving here, so im just using the original value
ggplot(gathered_continent_df, 
       aes(x=Year, y=value, fill = Continent)) +
      geom_bar(stat="identity") +
      scale_y_continuous(labels = comma)
```

```{r Freq Poly, everyone loves...no we dont..}
 #Freq Poly of forest area
ggplot(gathered_continent_df, aes(value,colour=Continent)) +
      geom_freqpoly(aes(group = Continent)) + scale_x_log10() +
      scale_x_continuous(labels = comma)
```


```{r}
#Max Value for each Continent
gathered_continent_df %>% 
      group_by(Continent) %>%
      slice(which.max(value)) 
```


Still Here? Please Submit some feedback, i will be happy if you suggest better graphics and techniques to do it better with large Datasets :), feel free to correct some grammar(sorry for my bad english :/) or bad code
